---
title: "ggplot2_1"
author: "Schwartz"
date: "9/19/2020"
output:
     html_document:
          toc: TRUE
          toc_depth: 4
          toc_float: true
          toc_collapsed: true
          theme: journal
          code_folding: hide
---

<!--- Change font size for headers --->
<style>
h1.title {
  font-size: 28px;
}
h1 {
  font-size: 22px;
}
h2 {
  font-size: 18px;
}
h3 { 
  font-size: 14px;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, 
                      fig.align = "center",
                      out.width = "80%")
```



# HW [Group Assignment]

*Each group member must uniquely complete this assignment for the group to get 
credit for this homework. If only some members of the group complete the 
assignment, they must each submit a supplemental explanation along with their 
submission as to why the other students in their group have not completed this 
assignment.*

0. Review the the 
   [ggplot2 documentation](https://r4ds.had.co.nz/data-visualisation.html)
   and related resources to your own satisfaction and as needed to create the 
   exercise
1. Find a data set you're interested in that lends itself to this assignment
    - As noted in `Pirates_Snakes.Rmd` and `Pirates_Snakes.ipynb` (from Class 5) 
      and in the code below (Class 6) some ways to find data are
      - data analysis websites such as [kaggle](https://www.kaggle.com/) data 
        repositories such as those provided by 
        [seaborn](https://github.com/mwaskom/seaborn-data) and
        [CMU](http://lib.stat.cmu.edu/datasets/) 
        (e.g., we've used the [penguins](https://raw.githubusercontent.com/mwaskom/seaborn-data/master/penguins.csv), 
         [strikes](http://lib.stat.cmu.edu/datasets/strikes), and
         [S&P](http://lib.stat.cmu.edu/datasets/spdc2693)) data sets)
      - other similar resources, such as (in no particular order)
        - [Data Portals](http://dataportals.org/)
        - [UCIML Repo](http://archive.ics.uci.edu/ml/)
        - [OpenDataMonitor](http://opendatamonitor.eu/)
        - [Quandl](http://quandl.com/)
        - [Wikipedia](https://en.wikipedia.org/wiki/List_of_datasets_for_machine-learning_research)
        - [AWS data set](https://registry.opendata.aws/)
        - [Quora.com](https://www.quora.com/Where-can-I-find-large-datasets-open-to-the-public)
        - [The datasets subreddit](https://www.reddit.com/r/datasets)
        - [TFDS](https://www.tensorflow.org/datasets/catalog/overview)
        
    - `data()` will show you data sets available in R
      - And you can load data as done in  `Pirates_Snakes.Rmd` (Class 6), 
        e.g., `data("USArrests")`
        - **HOWEVER**, if you use a built in data set, the maximum grade you can 
          get on this assignment is 90%, with the following exception:
          - 10% of the assignment is for "reading in" data from a `.csv` or other 
          tabular "non built-in" data source
          - ...so you you cat get full 100% by first saving an internal data set 
          to a `.csv` or other tabular data format and THEN reading it into R 
          (to get credit for the last 10%
    
2. Create an attractive graphical image presenting information in the data
3. Use both the `size` and `color` attributes in your graphic
    - These aesthetics are demonstrated in the "`hwy` versus `displ`" plots 
      [here](https://r4ds.had.co.nz/data-visualisation.html#aesthetic-mappings):
      - `ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))`
           # (colored by `class`)
      - `ggplot(data = mpg) +  geom_point(mapping = aes(x = displ, y = hwy, size = class))`    
           # (sized by `class`)

4. Use the `facets` feature of ggplot2
    - This aesthetic is demonstrated 
      [here](https://r4ds.had.co.nz/data-visualisation.html#facets)
5. For +10 points, use a "color blind friendly" color palette.
    - One example of how to do this with `scale_colour_manual(values=cbbPalette)` and
      `cbbPalette <- c("#009E73", "#CC79A7","#D55E00", "#000000","#56B4E9", "#F0E442", "#0072B2", "#E69F00")`
      is described [here](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette)

Hint: you'll need 3 factor variables and two continuous variables to do this, though continuous variables can be discretized to produce factor level variables.


# ggplot2 (R)

- Perhaps the best (most principled) plotting library around
- Here are the references/resources I've used to create the code in this file

## Canonical Resources:
- [R4DS -- Data Visualization](https://r4ds.had.co.nz/data-visualisation.html)
- [tiidyverse -- learning ggplot2](https://ggplot2.tidyverse.org/#learning-ggplot2)
- [R4DS -- Graphics for Communication](https://r4ds.had.co.nz/graphics-for-communication.html)
  - [R4DS -- scales](https://r4ds.had.co.nz/graphics-for-communication.html#scales)
    - [R4DS -- rescaling](https://r4ds.had.co.nz/graphics-for-communication.html#replacing-a-scale)
- [R Cookbook](http://www.cookbook-r.com/Graphs/)
- [tidyverse -- ggplot2](https://ggplot2.tidyverse.org)

## Manual Pages
- [tidyverse readr](https://readr.tidyverse.org/reference/read_delim.html)
- [tidyverse str_replace](https://stringr.tidyverse.org/reference/str_replace.html)
- [tidyverse join](https://dplyr.tidyverse.org/reference/join.html)
- [tidyverse scale_viridis](https://ggplot2.tidyverse.org/reference/scale_viridis.html)

## Additional Resources
- [Proportional Ink](https://stackoverflow.com/questions/7714677/scatterplot-with-too-many-points)
- [Rotatinig Labels](https://community.rstudio.com/t/x-axis-labels-overlap-want-to-rotate-labels-45/63800)
- [Customizinig Ticks](http://www.sthda.com/english/wiki/ggplot2-axis-ticks-a-guide-to-customize-tick-marks-and-labels)

## Data Sets
- [Kaggle 2016 Primaries](https://www.kaggle.com/benhamner/2016-us-election?select=primary_results.csv)
- [Kaggle County Facts](https://www.kaggle.com/benhamner/2016-us-election?select=county_facts.csv)

# Code

## Initialization

- loading tidyverse, data, and joining tables

```{r data}
library(tidyverse)

# https://readr.tidyverse.org/reference/read_delim.html
d1 <- "/Users/gck8gd/Documents/courses/SDS_3003_Communication_w_Data/primary_results.csv"
primary_results <- readr::read_csv(d1)
d2 <- "/Users/gck8gd/Documents/courses/SDS_3003_Communication_w_Data/county_facts.csv"
county_facts <- readr::read_csv(d2)

# https://stringr.tidyverse.org/reference/str_replace.html
# https://dplyr.tidyverse.org/reference/join.html
county_facts %>% dplyr::mutate(county=stringr::str_replace(area_name, 
                                                          ' County', '')) %>%
  dplyr::inner_join(primary_results, by=c('state_abbreviation','county')) ->
  county_factsANDprimaries
```

## Elements of a ggplot

### `ggplot` function

```{r components_1}
my_first_ggplot <- ggplot(data=county_factsANDprimaries)

my_first_ggplot
```

### `ggplot` parts

```{r components_2a}
ls(my_first_ggplot)
```

### `ggplot` data 

```{r components_2b}
my_first_ggplot[['data']]
```

### `aes` aesthetic mappings

```{r components_2c}
my_first_ggplot[['mapping']]
```

```{r components_2e}
my_first_ggplot <- ggplot(data=county_factsANDprimaries,
                          mapping=aes(x=PST045214, y=fraction_votes))

my_first_ggplot[['mapping']]
my_first_ggplot
```

### `geom` rendering 

```{r components_3}

my_first_ggplot + geom_point(mapping=aes(x=PST045214, y=fraction_votes)) -> 
  my_first_rendered_ggplot

my_first_rendered_ggplot
```

### adjusting scales

- show data on appropriate scales

```{r axes}
# https://r4ds.had.co.nz/graphics-for-communication.html#scales
# https://r4ds.had.co.nz/graphics-for-communication.html#replacing-a-scale
my_first_rendered_ggplot + scale_x_log10()
```


### rendering style type

- rule of proportional ink
- plotting etiquette: color vision deficiency

```{r density}
# https://stackoverflow.com/questions/7714677/scatterplot-with-too-many-points
my_first_ggplot + geom_point() + scale_x_log10()
my_first_ggplot + geom_point(alpha=0.1) + scale_x_log10()
my_first_ggplot + geom_point(alpha=0.1, size=2.5, stroke=0) + scale_x_log10()
my_first_ggplot + geom_hex() + scale_x_log10()
my_first_ggplot + scale_x_log10() + geom_density_2d()
my_first_ggplot + geom_hex() + geom_density_2d() + scale_x_log10()
# https://ggplot2.tidyverse.org/reference/scale_viridis.html
my_first_ggplot + geom_hex() + geom_density_2d() + scale_x_log10() +
  scale_fill_viridis_c()
my_first_ggplot + scale_fill_viridis_c() + scale_x_log10() +
  stat_density_2d(aes(fill=stat(density)), geom='raster', contour=FALSE) +
  coord_cartesian(expand=FALSE) + 
  geom_point(color='white', shape='.', size=0.0001, alpha=0.1)
```


### more `aes` options
- size
- color
- alpha

```{r aes}

my_second_ggplot <- ggplot(data=county_factsANDprimaries,
                           mapping=aes(size=POP060210, alpha=POP060210,
                                       color=party, 
                                       x=PST045214, y=fraction_votes))

my_second_ggplot + geom_point() + scale_x_log10()

```

## labels

- you will need to do something like this for your homework

```{r labels}

county_factsANDprimaries <- county_factsANDprimaries %>% 
  dplyr::rename(`County Population`=PST045214, 
                `County Density (per square mile)`=POP060210,
                Party=party, `Fraction of Votes`=fraction_votes) 


my_second_ggplot <- ggplot(data=county_factsANDprimaries,
                           mapping=aes(size=`County Density (per square mile)`, 
                                       alpha=`County Density (per square mile)`,
                                       color=Party, 
                                       x=`County Population`, 
                                       y=`Fraction of Votes`))

my_second_ggplot + geom_point() + scale_x_log10()
```

## facets

- this is exactly like what you will need to do for your homework
- but critique what I've done here
  - what do you like about it?
  - what don't you like about it?

```{r facets}

county_factsANDprimaries <- county_factsANDprimaries %>% 
  dplyr::mutate(Candidate = stringr::str_replace(candidate, ' ','\n'))

  
my_third_ggplot <- ggplot(data=county_factsANDprimaries,
                          mapping=aes(size=`County Population`, 
                                      alpha=`County Population`,
                                      color=Party, 
                                      x=`County Density (per square mile)`, 
                                      y=`Fraction of Votes`))

my_third_ggplot + geom_point() + scale_x_log10() +
  facet_wrap(~ Candidate, nrow = 2)
```

### touchups

```{r touchup}
# https://community.rstudio.com/t/x-axis-labels-overlap-want-to-rotate-labels-45/63800
# http://www.sthda.com/english/wiki/ggplot2-axis-ticks-a-guide-to-customize-tick-marks-and-labels
my_third_ggplot + geom_point() + scale_x_log10() +
  facet_wrap(~ Candidate, nrow=4, scales='fixed') + 
  scale_color_hue(direction=-1) +
  scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8, 1.0),
                     guide=guide_axis(n.dodge=2, check.overlap=TRUE)) +
  coord_flip()
```


